---
title: Personnal coding rules in R
author: Virgile
date: '2017-08-24'
slug: personnal-coding-rules-in-r
categories:
  - R
tags:
  - R
  - code
---


*Note: in the following doc, many R's error messages are displayed in French*

## References

[*The R Inferno* by Patrick Burns](http://www.burns-stat.com/pages/Tutor/R_inferno.pdf)

[Hadley's style guide](http://adv-r.had.co.nz/Style.html)

[Google's R style guide](https://google.github.io/styleguide/Rguide.xml)


## Style guide

Two references are unavoidable, the [Hadley's style guide](http://adv-r.had.co.nz/Style.html) and the [Google's R style guide](https://google.github.io/styleguide/Rguide.xml).

Has Hadley wrote, the most important is to have a consistent style, and clear.

In the following document, I use:


```{r, eval=FALSE}
# GOOD
variable.Name
variableName
variable_Name

# BAD
variablename
```


I also like to write the name of the variable in fist and then it's specification:

```{r, eval=FALSE}
# GOOD
kk = 1:5
kk_max = max(kk)
kk_sd = sd(kk)
kk_log10Mean = log10(mean(kk))

km = 1:10
km_max = max(km)
km_sd = sd(km)
km_log10Mean = log10(mean(km))

# BAD
max_kk = max(kk)
sd_kk = sd(kk)
log10Mean_kk = log10(mean(kk))

max_km = max(km)
sd_km = sd(km)
log10Mean_km = log10(mean(km))
```


## Don't do

Some letters are use in `R`, like `T` use for the boolean `TRUE`, therefore, do not use as variable:


```{r, eval=FALSE}
# BAD
T # TRUE
F # FALSE
TRUE
FALSE

# GOOD
Temperature ; Total
Fox ; Flower
```

Function already use:

```{r, eval=FALSE}
# BAD
c() # vector
lm() # linear model
```

## `<-` or `=`

Big question in R:
[revolutionanalytics?](http://blog.revolutionanalytics.com/2008/12/use-equals-or-arrow-for-assignment.html)
[stackoverflow](http://stackoverflow.com/questions/1741820/assignment-operators-in-r-and)

Let's take this function: `median(x, na.rm = FALSE, ...)`, and look at this:

```{r, eval=FALSE}
median(x <- 1:10)
median(x = 1:10)

# but:
median(a <- 1:10)
median(a = 1:10)

# and also:
b
median(b <- 1:10)
b
```


## Loop

```{r, eval=FALSE}
# GOOD
if (y == 0) {
  log(x)
} else {
  y ^ x
}

# BAD
if (y == 0) {
  log(x)
}
else {
  y ^ x
}
```


## Functions

- Deprecated: `deprecate()`
- Errors: `stop()`
- Warnings: `warning()`
- Return: `return()`


## Boolean

[From this answer in stackexchange](http://stackoverflow.com/questions/6558921/r-boolean-operators-and)

Difference between `&` and `&&`

```{r}
((-2:2) >= 0) & ((-2:2) <= 0)
((-2:2) >= 0) && ((-2:2) <= 0)
```


Difference between `|` and `||`

```{r}
((-5:-1) >= 0) | ((-2:2) <= 0)
((-5:-1) >= 0) || ((-2:2) <= 0)
```


```{r, eval=FALSE}
a
# Erreur : objet 'a' introuvable
TRUE || a
# [1] TRUE
FALSE && a
# [1] FALSE
TRUE | a
# Erreur : objet 'a' introuvable
FALSE & a
# Erreur : objet 'a' introuvable
```

